### 2019-11-05 latex/acrobot.tex
Changed the bibliography style This new version changes the bibliography type to use a different package, and we now compile with lualatex. It helps with error and warning messages, as it actually shows where the message is coming from for proper errors (as opposed to always doing it in main)
Added modelines to all my tex files
Filled out the first draft of acrobot.tex

%/===============/%
\section{Energy Injection for the Acrobot}

>>>>

%! TEX root = main.tex

%/========== Acrobot ==========/%
\section{Energy Injection for the Acrobot}
Suppose the acrobot is simple; 
that is, assume that the acrobot's two links are of equal length \(l\)
and of equal mass \(m\), with the mass concentrated at the tip of each rod.

Let \(q = (q_u,q_a)\) where \(q_u\) is the shoulder angle and \(q_a\) is the hip angle.

Then the acrobot has hamiltonian given by
\[
    \mathcal{H}(q,p) = \frac{1}{2} p^T M^{-1}(q) p + V(q)
\]
where
\[
    M^{-1}(q) = \frac{1}{m l^2 (2 - \cos^2(q_a))} 
    \begin{bmatrix}
        1              & -(1 + \cos(q_a)) \\
        -(1 + \cos(q_a)) & (3 + 2\cos(q_a))
    \end{bmatrix}
\]
and
\[
    V(q) = m g l (3 - 2 \cos(q_u) - \cos(q_u + q_a))
\]

Notice that the acrobot satisfies \(\partial M^{-1}(q)/ \partial q_u = 0\),
which means it may be possible to enforce a regular solvable VNHC to inject energy
into the acrobot.
One such VNHC is
\begin{align}\label{eqn:acrobot_vnhc}
    \begin{split}
    h(q,p) &= q_a - f(q_u,p_u) \\
           &= q_a - \bar{q_a}\sin(\arctan(p_u,q_u)) \\
           &= q_a - \frac{\bar{q_a}p_u}{\sqrt{q_u^2 + p_u^2}}
    \end{split}
\end{align}
where \(\bar{q_a} \in ]0,\pi[\) is a parameter which dictates how much the 
controlled link can swing back and forth.

To compute the regularity and solvability conditions, first observe that since
the acrobot has two coordinates and only \(p_a\) is actuatable, the control matrix
\(B = [0, 1]^T\) is constant and full rank everywhere. This also implies that
\(h\) is both regular and solvable if \(dh_q M^{-1}(q)B\) is full rank.

Observe that
\begin{align}
    \begin{split}
        dh_q M^{-1}(q)B &= \frac{1}{m l^2 (2 - \cos^2(q_a))}
\left[ \frac{\bar{q_a}p_uq_u}{(q_u^2 + p_u^2)^{3/2}}, 1\right] 
    \begin{bmatrix}
        -(1 + \cos(q_a)) \\
        (3 + 2\cos(q_a))
    \end{bmatrix} \\
                     &= \frac{ (3+2\cos(q_a))(q_u^2 + p_u^2)^{3/2} 
                     - \bar{q_a}p_uq_u(1+cos(q_a))}
                     {m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))}
    \end{split}
\end{align}

TODO: \textbf{Determine if this is full rank everywhere except at \((q_u,p_u) = (0,0)\)}

Solving for \(p_a\) is straightforward. The computation is shown below.
\begin{multline*}
    p_a = \left(dh_q M^{-1}(q)B\right)^{-1}  \\
        \left(\frac{m g l \bar{q_a}p_u q_u^2}{(q_u^2+p_u^2)^{3/2}} 
            \left(2\sin(q_u) + sin(q_u + q_a) \right)
        +
            \frac{1}{m l^2 (2 - \cos^2(q_a))}
            [ \frac{\bar{q_a}p_uq_u}{(q_u^2+p_u^2)^{3/2}}, 1] 
            \begin{bmatrix}
              1 \\
              -(1 + \cos(q_a))
          \end{bmatrix} \right)
\end{multline*}
\begin{align*}
    &= -\frac
    {
        \frac{ m g l \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))}{(q_u^2+p_u^2)^{3/2}}
        + \frac{\bar{q_a} p_u^2 q_u}{m l^2 (q_u^2 + p_u^2)^{3/2} (2 -\cos^2(q_a))}
        - \frac{(1+\cos(q_a)) p_u}{m l^2 (2 - \cos^2(q_a))}
    }
    {
        \frac
        {
            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
        }
        {
            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
        }
    } \\
    &= \frac
    {
        \frac
        {
            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
            - \bar{q_a}p_u^2 q_u 
            + (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
        }
        {
            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
        }
    }
    {
        \frac
        {
            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
        }
        {
            m l^2 (q_u^2+p_u^2)^{3/2} (2 - \cos^2(q_a))
        }
    }
\end{align*}

\begin{equation*}
    \Rightarrow p_a = \frac
        {
            (1 + \cos(q_a))(q_u^2 + p_u^2)^{3/2} p_u
            -m^2 g l^3 \bar{q_a}q_u^2(2\sin(q_u) + \sin(q_u+q_a))(2 - \cos^2(q_a))
            - \bar{q_a}q_up_u^2
        }
        {
            (3 + 2\cos(q_a))(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + \cos(q_a))q_up_u
        }
\end{equation*}

With \(p_a\) in hand, it is now possible to solve for the zero dynamics under the constraint. To simplify the notation we define \(s_u := \sin(q_u)\),
\(s_{ua} := \sin(q_u + q_a)\), 
and \(c_a := \cos(q_a)\).
\begin{align*}
    \dot{q_u} &= \frac{1}{m l^2 (2 - c_a^2)} [ 1, -(1+c_a)]
    \begin{bmatrix} p_u \\ p_a \end{bmatrix} \\
    \dot{p_u} &= - m g l (2 s_u + s_{ua})
\end{align*}
Expanding out \(\dot{q_u}\) we get
\begin{multline*}
    \dot{q_u} = 
    \frac{p_u}{m l^2 (2 - c_a^2)}\\
    + \frac
    {
        m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
        + \bar{q_a}(1+c_a)q_up_u^2
    }
    {
        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
    }
\end{multline*}
and, putting everything over one denominator,
\begin{align*}
    \dot{q_a} &= \frac
    {
        (3 + 2c_a)(q_u^2+p_u^2)^{3/2}p_u 
        - (1+c_a)^2(q_u^2+p_u^2)^{3/2}p_u 
        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(2 - c_a^2)(1 + c_a)
    }
    {
        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
    } \\
    &= \frac
    {
        (2-c_a^2)\left((q_u^2+p_u^2)^{3/2}p_u 
        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)\right)
    }
    {
        ml^2 (2-c_a^2) ((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u)
    }
\end{align*}
Therefore, the zero-dynamics of the acrobot under the constraint 
\(q_a = \sin(\arctan(p_u,q_u))\) are:
\begin{align}\label{eqn:acrobot_eom_qp}
\begin{split}
    \dot{q_u} &= \frac
    {
        (q_u^2+p_u^2)^{3/2}p_u 
        +m^2 g l^3 \bar{q_a}q_u^2(2s_u + s_{ua})(1 + c_a)
    }
    {
        ml^2 \left((3 + 2c_a)(q_u^2+p_u^2)^{3/2} - \bar{q_a}(1 + c_a)q_up_u\right)
    } \\
    \dot{p_u} &= - m g l (2s_u + s_{ua})
\end{split}
\end{align}

It may be convenient to represent these equations of motion  in polar coordinates.
Letting \(r = \sqrt{q_u^2 + p_u^2}\) be the radius and 
\(\theta = \arctan(p_u,q_u)\) be the angle in the \((q_u,p_u)\) plane and taking
derivatives, it is easy to show that
\begin{align}\label{eqn:dr_and_dt_general}
\begin{split}
    \dot{r} &= \sin(\theta)]
    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix} \\
    \dot{\theta} &= \frac{1}{r}[-\sin(\theta), \cos(\theta)]
    \begin{bmatrix} \dot{q_u}(r,\theta) \\ \dot{p_u}(r,\theta) \end{bmatrix}
\end{split}
\end{align}

By expanding out (\ref{eqn:dr_and_dt_general}) then substituting \(q_u = r\cos(\theta)\)
and \(p_u = r\sin(\theta)\)
it is straightforward computation to find the equations of motion in cylindrical \((r,\theta)\)
coordinates. To simplify the notation yet again, define
\(s(g)_\theta := \sin(g(\theta))\) and \(c(g)_\theta := \cos(g(\theta))\) 
for any function \(g: \mathbb{R} \rightarrow \mathbb{R}\). The final equations
of motion in \((r,\theta)\) coordinates are
\begin{align}\label{eqn:acrobot_eom_rt}
\begin{split}
    \dot{r} &= \frac
    {
        r^2 c_\theta s_\theta 
        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
        \left(
            (1+c(\bar{q_a}s)_\theta)\bar{q_a}c_\theta 
            - (3 + 2c(\bar{q_a}s)_\theta)s_\theta r
        \right)
    }
    {
        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
    }\\
    \dot{\theta} &= -\frac
    {
        r s_\theta^2
        + m^2 g l^3 (2 s(rc)_\theta + s(rc + \bar{q_a}s)_\theta)
        (3 + 2c(\bar{q_a}s)_\theta)c_\theta
    }
    {
        m l^2 \left((3+2c(\bar{q_a}s)_\theta)r 
        - \bar{q_a}(1+ c(\bar{q_a}s)_\theta)c_\theta s_\theta\right)
    }
\end{split}
\end{align}

%/========== /Acrobot ==========/%
% vim: set ts=4 sw=4 sts=0 et ffs=unix :
